Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)

### 2014-04-28 manuscript.tex
edits
text

Given that the Hopscotch insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte. However, in sampling numerous individuals from many teosinte populations our study provides greater insight into the distribution and prevalence of the Hopscotch in teosinte. While our findings are consistent with a previous study by \citet{Studer2011} in that we identified the Hopscotch allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}. Many of our populations with high frequency of the Hopscotch allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus than in the Balsas region where maize was domesticated \jri{cite matsuoka} . While gene flow from crops into their wild relatives is well-known, (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}),\jri{do all these show evidence of introgression AT domestication loci, or just introgression in general?}  our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}. Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize) \citep{Hufford2013}.

@@

Other explanation for differential frequencies of the Hopscotch among teosinte populations include both drift and natural selection.  Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyh\"aj\"arvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings. \jri{deviations from HWE may be consistent too if we see excess of homozygotes. do we?}. This bottleneck, however, does not explain differences in Hopscotch allele frequency among populations, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyh\"aj\"arvi2013} is not suggestive of colonization or other demographic events that might predict a high frequency of the allele in multiple populations.  \jri{here we need a few sentences on selection.  the 5' UTR has much more negative D than the upstream. do we know the Hop genotype for sequenced lines? can we separate the sequences into hop/no hop and look for differences? it wasn't until we did this that gt1 stuff really popped out.}
%   from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..

>>>>

Given that the Hopscotch insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte. However, in sampling numerous individuals from many teosinte populations our study provides greater insight into the distribution and prevalence of the Hopscotch in teosinte. While our findings are consistent with a previous study by \citet{Studer2011} in that we identified the Hopscotch allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}. Many of our populations with high frequency of the Hopscotch allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus than in the Balsas region where maize was domesticated \citep{Matsuoka2002}. While gene flow from crops into their wild relatives is well-known, (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}),\jri{do all these show evidence of introgression AT domestication loci, or just introgression in general?}  \lev{just introgression in general, I believe Thurber and Huff looked at specific domestication loci} our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}. Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize) \citep{Hufford2013}.

@@

Other explanation for differential frequencies of the Hopscotch among teosinte populations include both drift and natural selection.  Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyh\"aj\"arvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings. \jri{deviations from HWE may be consistent too if we see excess of homozygotes. do we?}. This bottleneck, however, does not explain differences in Hopscotch allele frequency among populations, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyh\"aj\"arvi2013} is not suggestive of colonization or other demographic events that might predict a high frequency of the allele in multiple populations.  \jri{here we need a few sentences on selection.  the 5' UTR has much more negative D than the upstream. do we know the Hop genotype for sequenced lines? can we separate the sequences into hop/no hop and look for differences? it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check}
%   from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..
